### 2014-02-26 progress-report/report.tex

\documentclass{article}
\usepackage{graphicx}
\usepackage{comment}
\usepackage{natbib}
\usepackage{listings}
\lstset{
    language=C,
    emph={nn_send, nn_receive, nn_configure, nn_invoke},
    emphstyle=\bfseries,
    basicstyle=\ttfamily\footnotesize,
    aboveskip=1mm plus 1mm minus 1mm,
    belowskip=1mm plus 1mm minus 1mm,
    mathescape=true,
    xleftmargin=\parindent,
}

\newcommand{\code}{\lstinline[basicstyle=\ttfamily\small]}

\newcommand{\todo}[2]{
\ifx\PrintTodo\undefined
\textcolor{red}{#2\xspace}
\else
\textcolor{red}{TODO[#1]:} \textcolor{blue}{\textit{#2}\xspace}
\fi
}

\begin{document}

\title{Visualizing Physical Query Execution in a Relational Big Data Management System}
\author{Umar Javed, Thierry Moreau, Dominik Moritz, Adriana Szekeres}

\maketitle

\input{related_work}
\input{project_plan}

\bibliographystyle{plain}
\bibliography{report}

\end{document}