### 2014-02-26 progress-report/project_plan.tex

\section{Project Plan}

Our team of four is led by Dominik, who started working on a query visualization framework for Myria last quarter. The responsibilities for each of the group members are listed below:

\begin {itemize}
    \item Umar Javed: visualizing network data flow between workers
    \item Thierry Moreau: visualizing query plan
    \item Dominik Moritz: team leader, in charge of data collection and back end
    \item Adriana Szerkes: visualizing query fragment execution
\end{itemize}