### 2014-02-26 progress-report/project_plan.tex
Added project milestones

\section{Project Plan}

\subsection{Team Responsibilities}

Our team of four is led by Dominik, who started working on a query visualization framework for Myria last quarter. The responsibilities for each of the group members are listed below:

\begin {itemize}
    \item Umar Javed: visualizing network data flow between workers
    \item Thierry Moreau: visualizing query plan
    \item Dominik Moritz: architect, data collection and back end
    \item Adriana Szerkes: visualizing query fragment execution
\end{itemize}

\subsection{Milestones}

\paragraph{Current Prototype}

\paragraph{Milestones}

\begin{enumerate}
    \item Project Planning (completed Feb 23rd)
    \begin{enumerate}
        \item Brainstorming and goals
        \item List questions we'd like to answer
        \item Develop a consistent story-board
        \item Assign responsibilities
    \end{enumerate}
    \item Initial System Setup \& Specifications (completed Feb 23rd)
    \begin{enumerate}
        \item Set up cluster \& web front-end
        \item Data collection from query-runs
        \item Formalize interfaces and data formats
    \end{enumerate}

\end{enumerate}