### 2014-02-26 progress-report/project_plan.tex
Add paragraph about current prototype.

\section{Project Plan}

\subsection{Team Responsibilities}

Our team of four is led by Dominik, who started working on a query visualization framework for Myria last quarter. The responsibilities for each of the group members are listed below:

\begin {itemize}
    \item Umar Javed: visualizing network data flow between workers
    \item Thierry Moreau: visualizing query plan
    \item Dominik Moritz: architect, data collection and back end
    \item Adriana Szerkes: visualizing query fragment execution
\end{itemize}

\subsection{Milestones}

\paragraph{Current Prototype}

Dominik built a prototype last quarter for the \emph{Distributed Systems} class. The goal of this project was to build a system to collect and analyze logs in the Myria \footnote{\url{http://myria.cs.washington.edu/}} system. The prototype allows a user to switch between three levels of query execution visualization to identify the problem at the granularity that best reveals the information they need. The goal for this project is to improve the visualization part of the previous prototype and build a more consistent exploration tool. Furthermore, the back end has to be rewritten to support visualization of long running queries in the Myria system but are not supported in the current prototype.

\paragraph{Milestones}

\begin{enumerate}
    \item Project Planning (completed Feb 23rd)
    \begin{enumerate}
        \item Brainstorming and goals
        \item List questions we'd like to answer
        \item Develop a consistent story-board
        \item Assign responsibilities
    \end{enumerate}
    \item Initial System Setup \& Specifications (completed Feb 23rd)
    \begin{enumerate}
        \item Set up cluster \& web front-end
        \item Data collection from query-runs
        \item Formalize interfaces and data formats
    \end{enumerate}

\end{enumerate}